Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-10-08 exp1/present2013.tex
full expirment now ok
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@654 43ef1190-90fd-4dd7-a769-d0121f02b80e

  \end{itemize}
\end{frame}
\note{Implemented in R, using packages \texttt{DoE.base} and \texttt{FrF2}.}
\note{Two experiment machines, one machine running queries}
\note{No warm-up, no cache busting}

@@

\end{frame}
% TODO explain Lenth
\note{Don't worry about the axes}
\note{Yes, the Implementation is better, we knew that, joins are
  naturally strong}
\note{``Union'', ``Optional'' and ``Lang'' demanding things}

\begin{frame}
\begin{table}[ht!]
\begin{center}
\caption{$p$-values for different parts of the experiment}\label{tab:pvaluesfull}
\begin{tabular}{cccccl}
  \hline
``TripleC'' & ``BGPComp'' & ``Lang'' & ``Union'' & ``Optional'' & $p$ \\ 
  \hline
  1 & 1 & 1 & 1 & 1 & 0.012 \\ 
  1 & 1 & 1 & 1 & 2 & $1.4 \cdot 10^{-09}$ \\ 
  1 & 1 & 1 & 2 & 1 & $3.1 \cdot 10^{-09}$ \\ 
  1 & 1 & 1 & 2 & 2 & $6.1 \cdot 10^{-11}$ \\ 
  1 & 1 & 2 & 1 & 1 & $3.3 \cdot 10^{-06}$ \\ 
  1 & 1 & 2 & 1 & 2 & $2.7 \cdot 10^{-09}$ \\ 
  1 & 1 & 2 & 2 & 1 & $2 \cdot 10^{-06}$ \\ 
  1 & 1 & 2 & 2 & 2 & $3.6 \cdot 10^{-10}$ \\ 
  1 & 2 & 1 & 1 & 1 & 0.014 \\ 
  1 & 2 & 1 & 1 & 2 & $1.2 \cdot 10^{-10}$ \\ 
  1 & 2 & 1 & 2 & 1 & $2.8 \cdot 10^{-14}$ \\ 
  1 & 2 & 1 & 2 & 2 & $4.1 \cdot 10^{-15}$ \\ 
  1 & 2 & 2 & 1 & 1 & $2.1 \cdot 10^{-05}$ \\ 
  1 & 2 & 2 & 1 & 2 & $2.7 \cdot 10^{-07}$ \\ 
  1 & 2 & 2 & 2 & 1 & 0.0072 \\ 
  1 & 2 & 2 & 2 & 2 & $1.6 \cdot 10^{-05}$ \\ 
  2 & 1 & 1 & 1 & 1 & 0.28 \\ 
  2 & 1 & 1 & 1 & 2 & $3 \cdot 10^{-07}$ \\ 
  2 & 1 & 1 & 2 & 1 & $3.3 \cdot 10^{-07}$ \\ 
  2 & 1 & 1 & 2 & 2 & $1.7 \cdot 10^{-08}$ \\ 
  2 & 1 & 2 & 1 & 1 & 0.0023 \\ 
  2 & 1 & 2 & 1 & 2 & $6.5 \cdot 10^{-07}$ \\ 
  2 & 1 & 2 & 2 & 1 & 0.00032 \\ 
  2 & 1 & 2 & 2 & 2 & $1.3 \cdot 10^{-06}$ \\ 
  2 & 2 & 1 & 1 & 1 & 0.013 \\ 
  2 & 2 & 1 & 1 & 2 & $1 \cdot 10^{-11}$ \\ 
  2 & 2 & 1 & 2 & 1 & $3.8 \cdot 10^{-11}$ \\ 
  2 & 2 & 1 & 2 & 2 & $4.1 \cdot 10^{-15}$ \\ 
  2 & 2 & 2 & 1 & 1 &   1 \\ 
  2 & 2 & 2 & 1 & 2 & $2.3 \cdot 10^{-05}$ \\ 
  2 & 2 & 2 & 2 & 1 &   1 \\ 
  2 & 2 & 2 & 2 & 2 & 0.99 \\ 
   \hline
\end{tabular}
\end{center}
\end{table}

>>>>

  \end{itemize}
\end{frame}
\note{Implemented in R, using packages \texttt{DoE.base} and \texttt{FrF2}.}
\note{Two experiment machines, one machine running queries}
\note{No warm-up, no cache busting}
\note{256 experiments in full factorial, no replications, compare
  favorably to benchmarks}

\subsection{Full Factorial Experiment}

@@

\end{frame}
% TODO explain Lenth
\note{Don't worry about the axes}
\note{Yes, the Implementation is better, we knew that, joins are
  naturally strong}
\note{``Union'', ``Optional'' and ``Lang'' demanding things}
\note{``BGPComp:Lang'' and ``Lang:Union'' interactions might be due to
  that many more triples need to be searched for a language tag in one
  of the levels of the interacting factors.}
\note{But what about ``Implement::TripleC''?}
\note{We want to do a hypothesis test}
\note{Two factors do not contribute significantly to the variation
  ``Range'' and ``Machine''}

\begin{frame}{Hypothesis tests}

Hypothesis formulation:
  \begin{itemize}
  \item $H_0$: The new implementation is no better than the old.
  \item $H_1$: The new implementation is better than the old.
  \end{itemize}

Classifying factors:
  \begin{itemize}
  \item ``Range'' and ``Machine'' are called \emph{inactive factors}.
  \item ``BGPComp'', ``Lang'', ``Optional'', ``Union'' and ``TripleC''
    are called \emph{environmental factors}.
  \item ``Implement'' is called a \emph{control factor}.
  \end{itemize}

\end{frame}
\begin{frame}{Hypothesis tests}

\begin{itemize}
\item Average over all the 32 level-combinations of the environmental
  factors.

% latex table generated in R 2.15.1 by xtable 1.5-6 package
% Tue Oct  8 14:51:28 2013
\begin{table}[ht]
\begin{center}
\begin{tabular}{rlll|r}
  \hline
 & Implement & Machine & Range & Response \\ 
  \hline
1 & 1 & 1 & 1 & 33.47 \\ 
  2 & 2 & 1 & 1 & 11.82 \\ 
  3 & 1 & 2 & 1 & 32.16 \\ 
  4 & 2 & 2 & 1 & 11.56 \\ 
  5 & 1 & 1 & 2 & 33.97 \\ 
  6 & 2 & 1 & 2 & 12.46 \\ 
  7 & 1 & 2 & 2 & 32.90 \\ 
  8 & 2 & 2 & 2 & 11.60 \\ 
   \hline
\end{tabular}
\end{center}
\end{table}

\item May create a one-sided two-sample t-test with 4 values for each of the levels.
\item Lends support to $H_1$ with a high probability, $p=1.16 \cdot
10^{-07}$.

\end{frame}
\note{The paper shows how similar testing can be done to investigate
  when the new implementation fails.} % TODO: Add more on this?
